### 2011-06-15 Spatial_adaptation/panmixia.tex
Undo revision 3397 by [[Special:Contributions/Peter|Peter]] ([[User talk:Peter|talk]])
/* Comparison to panmixia */
/* Comparison to panmixia */  small fix

The first question is something we did not include explicitly in our first paper.
The area occupied by an advantageous mutation grows quadratically, 
but only after a (random) initial period of local fixation.
We can call ``local fixation'' 
the time when the growth of a panmictic and a spatially restricted family
begins to differ significantly,
which occurs when the number of individuals is $\sigma^2 \rho$.
This time is thus approximately
\[
  T_l = \frac{1}{s} \log (\sigma^2 \rho) .
\]
(The standard deviation of this time is of the same order; let's ignore the randomness.)
On the other hand, the time we have to wait to see a new, successful mutation is on the order of
\[
  T_n = \frac{1}{4N\mu s} .
\]
Therefore, we will see new mutations in the time it takes a new mutation to fix locally
(and hence, the time it takes a wave to get going)
if
\[
    T_l > T_n ,
\]
or since $\rho = N/A$,
\begin{equation} \label{eqn:tlggtn}
  2 \log(\sigma / \sqrt{A}) + \log(N) > \frac{1}{4\mu N} .
\end{equation}
A reasonable range for $r=\sqrt{A}/\sigma$ is $10 \le r \le 1000$,
over which range $4 < 2\log(r) < 14$, so $\log(N)$ will dominate the inequality.
In other words, to get new mutations in the time it takes to locally fix,
we'd need $1/(4 N \log N) < \mu$.
Setting $\mu=10^{-8}$ tells us that we get multiple single base mutations in this time 
if $N$ is larger than about a million.

@@

% \begin{figure}[hht]
  \includegraphics[width=1.0\textwidth]{min-s-versus-r.pdf}
% \caption{Plot of $s= 8 \left( \log(r)/r \right)^2$ against $r=\sqrt{A/\sigma^2}$.} 
% \end{figure}

@@

\includegraphics[width=1.0\textwidth]{f-versus-gamma.pdf}

@@

\includegraphics[width=1.0\textwidth]{gamma-contour-1e-8.pdf}

>>>>

The first question is something we did not include explicitly in our first paper.
The area occupied by an advantageous mutation grows quadratically, 
but only after a (random) initial period of local fixation.
We can call ``local fixation'' 
the time when the growth of a panmictic and a spatially restricted family
begins to differ significantly,
which occurs when the number of individuals is $\sigma^2 \rho$.
This time is thus approximately
\[
  T_l = \frac{1}{s} \log (\sigma^2 \rho) .
\]
(The standard deviation of this time is of the same order; let's ignore the randomness.)
On the other hand, the time we have to wait to see a new, successful mutation is on the order of
\[
  T_n = \frac{1}{4N\mu s} .
\]
Therefore, we will see new mutations in the time it takes a new mutation to fix locally
(and hence, the time it takes a wave to get going)
if
\[
    T_l > T_n ,
\]
or since $\rho = N/A$,
\begin{equation} \label{eqn:tlggtn}
  2 \log(\sigma / \sqrt{A}) + \log(N) > \frac{1}{4\mu N} .
\end{equation}
A reasonable range for $r=\sqrt{A}/\sigma$ is $10 \le r \le 1000$,
over which range $4 < 2\log(r) < 14$, while perhaps $10 \le \log(N) \le 20$.
In other words, to get new mutations in the time it takes to locally fix,
we'd need $1/(4 N (\log N + 2 \log r)) < \mu$.
Setting $\mu=10^{-8}$ and taking $r$ in this range
tells us that we only get multiple single base mutations in this time 
if $N$ is larger than about a million.

@@

 \begin{figure}[hht]
  \includegraphics[width=0.8\textwidth]{min-s-versus-r.png}
 \caption{Plot of $s= 8 \left( \log(r)/r \right)^2$ against $r=\sqrt{A/\sigma^2}$.} 
 \end{figure}

@@

\includegraphics[width=1.0\textwidth]{f-versus-gamma}

@@

\includegraphics[width=0.8\textwidth]{gamma-contour-1e-8.png}
